---
title: "ph_1"
author: "Ryan Wang"
date: "7/2/2022"
output: html_document
---

```{r setup, include=FALSE, messsage = FALSE, warning= FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(here)
library(ggplot2)
library(lubridate)
library(scales)
library(car) #for the qqPlot function
library(psych) #for pairs.panels()
library(knitr)
library(nlme)
library(multcomp)
library(dplyr)
```

```{r}
# load the data
ph_<-read.csv('C:/Users/Ryan_/Documents/R/time series/files/ph_across_sites.csv')
ph1<-subset(ph_, select = -c(temp_durafet_c, temp_hobo_c)) 
ph_total<-subset(ph1,  ph1$tide == 'high') # filter the data
write.csv(ph_total,"C:/Users/Ryan_/Documents/R/time series/files/ph_total.csv", row.names = FALSE)

ph_total$site<-as.factor(ph_total$site)
levels(ph_total$site)

ph_subset<-ph_total[c(502:573, 12184: 12221, 17666: 17741), ]

```

```{r}
alg <- ph_ %>%
  filter(site=="Alegria")%>%
  mutate(date_time=mdy_hm(date_time))

lol <- ph_ %>%
  filter(site=="Lompoc Landing") %>%
  mutate(date_time=mdy_hm(date_time))

str(lol)
bml <- ph_ %>%
  filter(site=="Bodega Bay")%>%
  mutate(date_time=mdy_hm(date_time))

ph_  <- ph_ %>%
  mutate(date_time=mdy_hm(date_time))
str(bml)
str(ph_)
```


```{r}
ph_$site <- factor(ph_$site, levels=c("Lompoc Landing","Alegria","Bodega Bay"))
ggplot(bml, aes(x=date_time, y=temp_c)) +
  geom_point() +
  geom_smooth(method="loess", span=0.2) 

pal <- c(
  "Alegria" = "#CC6677",
  "Lompoc Landing" = "#44AA99",
  "Bodega Bay" = "#6699CC"
)
#Set x axis limits
lims <- as.POSIXct(strptime(c("2021-05-30 00:00:00", "2021-11-08 23:45:00"), 
                   format = "%Y-%m-%d %H:%M:%S"))

ggplot(lol, aes(x=date_time, y=p_h, group=site)) +
  geom_line(aes(color=site), size=0.7, alpha=0.4) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  geom_smooth(aes(color=site), method="loess", span=0.1) +
  scale_x_datetime(breaks = scales::date_breaks("4 weeks"), 
                   labels = date_format("%m/%d %H:%m")) +
  xlab("Date and time") +
  ylab("pH")


ggplot(ph_, aes(x=date_time, y=temp_c, group=site)) +
  geom_line(aes(color=site), size=0.7, alpha=0.4) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  geom_smooth(aes(color=site), method="loess", span=0.1) +
  xlab("Date and time") +
  ylab("Temperature") 
```

```{r}
str(ph_)
raw_combine <- ph_ %>%
  dplyr::select(site, date_time, temp_c, p_h) %>%
  pivot_longer(cols=temp_c:p_h,
               names_to = "group",
               values_to = "value")

ggplot(raw_combine, aes(x=date_time, y=value, group=group, color=site)) +
  #geom_line(aes(color=site), size=0.7, alpha=0.2) +
  geom_line(aes(x=date_time, y=value, group=group), data = . %>% filter((site %in% c("Alegria"))), alpha=0.4) + #plot only ALG line
  geom_line(aes(x=date_time, y=value, group=group), data = . %>% filter((site %in% c("Lompoc Landing"))), alpha=0.4) + #plot only LOL
  geom_line(aes(x=date_time, y=value, group=group), data = . %>% filter((site %in% c("Bodega Bay"))), alpha=0.4) + #plot only BML
  geom_smooth(aes(group=site), method="loess", span=0.1, se=FALSE) + 
  scale_color_manual(values = pal) + #color lines by custom site color palette
  #geom_point(aes(color=group), size=0.5) +
  scale_x_datetime(breaks = date_breaks("1 week"),
                   labels = date_format("%b-%d"),
                   limits = lims,
                   # limits = as.POSIXct(strptime(c("2021-06-20 07:00",
                   #                                "2021-10-08 07:00"),
                   #                              format = "%Y-%m-%d %H:%M")),
                    # c(as.POSIXct("2021-06-14 00:07:30"),as.POSIXct("2021-10-08 00:07:00")),
                     expand = c(0, 0)) + 
  facet_grid(group ~ ., #facet wrap to create one panel for pH and one for temp
             scales = "free_y",
             switch="both",
             labeller = as_labeller(c(temp_c = "Temperature (C)", p_h = "pH"))) + #customize strip labels
  xlab("Date and time") +
  ylab(NULL) + #remove "Value" from Y axis label
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=22),
        #axis.title.x=element_text(size=25),
        axis.title.x=element_blank(),
        axis.text.y=element_text(size=22),
        legend.text=element_text(size=20),
        legend.title = element_blank(),
        #element_text(size=25),
        legend.position = c(0.91,0.60), #customize legend position on plot (bottom right)
        #legend.position = c(0.92,0.95) #put legend on top right of plot
        #legend.key.size = unit(4,"line"),
        #legend.key.height = unit(1,"cm"),
        strip.background = element_blank(), #remove strip background from facet_grid
        strip.text.y = element_text(size = 25),
        strip.placement = "outside") + #place the strip outside of the plot
  guides(colour = guide_legend(override.aes = list(size=5)))
```

## Use one-way ANOVA to compare the mean
Null hypothesis: The means of the pH of the three sites are equal.
Alternative hypothesis: At least one means of the pH across the three sites are not equal to others.
```{r}
boxplot(ph_subset$p_h~ph_subset$site, xlab = 'Sites', ylab = 'pH') # visualize data

ph_aov<-aov(ph_subset$p_h~ph_subset$site, data = ph_subset)
par(mfrow = c(2,2))
plot(ph_aov)
leveneTest(ph_subset$p_h~ph_subset$site, data = ph_subset) # test equality of variance
# Even though the levene's test gives a low p-value, the residual vs. fitted plot shows no pattern and the predicted line is straight. So, I assume that the variances are the same

res_ph_aov <- ph_aov$residuals
qqPlot(res_ph_aov)
shapiro.test(res_ph_aov) # test normality

ph_subset$log_ph<-log(ph_subset$p_h)
log_ph_aov<-aov(ph_subset$log_ph~ph_subset$site, data = ph_subset)
par(mfrow = c(2,2))
plot(log_ph_aov)
leveneTest(ph_subset$log_ph~ph_subset$site, data = ph_subset) # test equality of variance
res_log_ph_aov <- log_ph_aov$residuals
qqPlot(res_log_ph_aov)
shapiro.test(res_log_ph_aov)

hist(ph_subset$p_h)
hist(ph_subset$log_ph)
# the residual does not look like normal even after transformation. But because the most of the points lie inside the 95% confidence interval and the sample size is large (central limit theorem applies), I assume that the residuals are normal.

summary(log_ph_aov)
TukeyHSD(log_ph_aov)
```
The p-value of the ANOVA is 2e-16, which is less than 0.05. 
So, I reject the null that the mean pH of site 1 = the mean pH of site 2 = the mean pH of site 3.
The Tukey-Kramer test shows that the mean pH at the three sites are significantly different from one another.

Do the same to the temperature.
```{r}
boxplot(ph_subset$temp_c~ph_subset$site, xlab = 'Sites', ylab = 'Temperature') # visualize data

temp_aov<-aov(ph_subset$temp_c~ph_subset$site, data = ph_subset)
par(mfrow = c(2,2))
plot(temp_aov)
leveneTest(ph_subset$p_h~ph_subset$site, data = ph_subset) 

res_temp_aov <- temp_aov$residuals
qqPlot(res_temp_aov)
shapiro.test(res_temp_aov) # not normal

kruskal.test(ph_subset$temp_c, ph_subset$site, data = ph_subset) # Non-parametric version of one-way ANOVA
install.packages('pgirmess')
library('pgirmess')
kruskalmc(ph_subset$temp_c, ph_subset$site, data = ph_subset) # # Non-parametric version of Tukey-Kramer test.
```
It shows that there exist significant temperature differences between Alegria-Bodega Bay and Bodega Bay-Lompoc Landing.

## Test correlation
```{r}
# 5:47 AM sunrise
# 8:14 PM sunset
ph_subset$sun<-"NO"
ph_subset[8:58, 'sun']<-"YES"
ph_subset[75:96, 'sun']<-"YES"
ph_subset[124:172, 'sun']<-"YES"

ph_l<-subset(ph_subset, select = c(site, p_h, temp_c, tide_height, sun)) # subset used for linear regression
pairs.panels(ph_l)
# The graph shows that the assumptions of bivariate normality are not met. 
# Log-transform the temperature and the tide height to see if it makes them more normal.
ph_l$log_temp<-log(ph_l$temp_c)
ph_l$log_tide_height<-log(ph_l$tide_height)
pairs.panels(ph_l)
# It shows that the pH and the log of tide height might be bivariate normally distributed.
# But the pH and the log of temperature are not.

cor.test(ph_l$p_h, ph_l$temp_c, method = 'spearman', alternative = 'two.sided') 
# Use Spearman's Rank Correlation because of no bivariate normal relationship.
# The p-value is 0.266, which is greater than the significant value, 0.05. I fail to reject the null that there exists correlation relationship between the temperature and the pH.

cor.test(ph_l$p_h, ph_l$log_tide_height, method = 'pearson', alternative = 'two.sided')
cor.test(ph_l$p_h, ph_l$tide_height, method = 'spearman', alternative = 'two.sided')
# Both Pearson's correlation test and Spearman's Rank Correlation show that there exists correlation between the pH and the log of tide height. The correlation coefficient is 0.082, which is very low.
```


## Use linear regression to predict the pH
## predictors: temperature and tide height (numeric); site and presence of sun (categorical).
```{r}
ph_null<-lm(p_h~1, data = ph_l)
anova(ph_null)
ph_onlyheight<-lm(p_h~tide_height, data = ph_l)
anova(ph_onlyheight)

ph_height_and_temp<-lm(p_h~tide_height+temp_c, data = ph_l)
anova(ph_height_and_temp)

ph_height_and_site<-lm(p_h~tide_height+site, data = ph_l)
anova(ph_height_and_site)

ph_height_and_sunlight<-lm(p_h~tide_height+sun, data = ph_l)
anova(ph_height_and_sunlight)

ph_height_sun_site<-lm(p_h~tide_height+sun+site, data = ph_l)
anova(ph_height_sun_site)
summary(ph_height_sun_site)

ph_full<-lm(p_h~tide_height+sun+site+temp_c, data = ph_l)
anova(ph_full)

ph_height_temp_site<-lm(p_h~tide_height+site+temp_c, data = ph_l)
anova(ph_height_temp_site)
summary(ph_height_temp_site)
AIC(ph_height_temp_site)
summary(ph_full)
summary(ph_height_and_temp)
# boxplot(ph_l$p_h~ph_l$sun)

# test the normality of residuals of the selected model. 
shapiro.test(residuals(ph_full)) # p-value = 0.00174. Reject the null that the residual is normal.
par(mfrow =c(2,2))
plot(ph_full)
# The shapiro-wilk test shows that the residuals are not normal.
# But the qqPlot shows that it is normal.
```
Predictors are tide height, site and presence of sun.


```{r}
result <- AIC(ph_null, ph_onlyheight, ph_height_and_temp, ph_height_and_site, ph_height_and_sunlight,ph_height_sun_site, ph_full) 

models <- list(ph_null, ph_onlyheight, ph_height_and_temp, ph_height_and_site, ph_height_and_sunlight, ph_height_sun_site, ph_full) 
result$BIC <- sapply(models, BIC) 

model_summary <- lapply(models, summary) 

for(i in 1:length(models)){ 
  result$rsq[i] <- model_summary[[i]]$r.squared 
  result$adj_rsq[i] <- model_summary[[i]]$adj.r.squared 
} 

kable(result, digits = 2, align = "c")
```

